Predicting Audio Features with Last.fm Tags

### 2023-05-02 paper/main.tex
first "readable" version of the intro

MIR applications often attempt to extract information from the music audio signal,
but analyzing associated metadata is also a common practice.
Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based representations.
Associated metadata, such as editorial information, lyrics, or user-generated content is usually in text format.
However, metadata can be available as images of videos, for example, when analyzing album artwork, or music videos.

@@

Many of the challenges in this field use audio features to predict other music-related aspects.
Examples of these features are 
For example, some studies have been using these features to predict the probably of a track being a hit [REF].

To make these predictions, researchers have been using machine learning models, specially over the past few years.
Previously, researchers commonly used hand-crafted audio computational models to perform these tasks.

But what if we tried to use these audio features as the predicted values? We would need a set of predictor variables to
acurrately predict the value of audio features.
This premise is the core concept of this article.

The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.

@@

>>>>

MIR applications often attempt to extract information from the music audio signal,
but analyzing associated metadata is also a common practice.
Audio signals are typically preprocessed and transformed into intermediate formats, such as frequency-based signal representations or hand-crafted audio features.
Associated metadata, such as editorial information, lyrics, or user-generated content is usually in text format.
However, metadata can be available as images of videos, for example, when analyzing album artwork, or music videos.

@@

Among potentially useful input and output values, research has proved Spotify audio features and Last.fm tags to be significant values to characterize music.
Spotify audio features capture high-level information about the music signal.
Examples of these values are energy, danceability, or valence, among others.
Last.fm tags are text labels that users associate to particular tracks, artists or albums in the Last.fm website.

Both types of values have been used as input values mostly for classification tasks, such as music genre recognition,
where, given Spotify features or Last.fm tasks, the model estimates the music genres of a particular track.
However, these values have not been used as target values by previous research, to the best of our knowledge.
In particular, the focus of this article is on predicting Spotify audio features, given a set of Last.fm tags.

By predicting Spotify audio features, we explore the relationship between subjective perception and concrete musical features.
This approach might help to identify patterns and hidden correlations between how music is percevied, consumed and discovered.

Additionally, predicted audio features could be used in recommendation systems to provide users with explainable recommendations.
Music recommendations,are not always easy to interpret from the perspective of the listener.
Users often get recommendations without meaningful explanations or justifications.
Therefore, by using predicted Spotify features as basis for recommendation, we can explain users why the algorithm suggests a particular track.
This process could be part of an explainable recommendation pipeline, where users enter a set of tags,
and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.

% The idea is, given a set of tags, to predict a set of audio features that a hypothetic track would exhibit.
% Then, we could build a track selection algorithm that selects actual tracks that are the closest to the predicted audio features.
% This process could be part of an explainable recommendation pipeline, where users enter a set of tags, and they recieve the predicted audio features, the closest tracks to those features, and the distance values between each track and the predicted features.

@@

In general, we believe that this novel approach that has the potential to benefit both listeners and researchers.
By combining subjective user-generated tags with objective audio features,
we can gain new insights into the complex relationship between perception and audio signal in music.

Our approach also presents limitations.
One limitation is the assumption of a strong relationship between Last.fm tags and Spotify features, which may not be true in all cases.
Future work could explore other sources of input values, possibly related to the user context, to improve the accuracy of the predictions.
